---
title: "FibroblastAnalysisincl.D8"
author: "A.DeMartin"
date: "2025-03-20"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages
```{r load packages}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Mm.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
```

# pre-processing 
## load files and merge
```{r load indiviual files, eval=FALSE, include=TRUE}
##load objects juli2021 and merge all 
basedir <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/dataJuli2021/"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratN")){
    seuratN <- merge(x = seuratN, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Total cells:", ncol(seuratN), "\n")
  }else{
    seuratN <- seuratS
    cat("Initialized with first Seurat object:", fileNamList[i], "\n")
  }
}
remove(seuratS)
unique(seuratN$dataset)

##load objects juli2022 and merge all
basedir <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/dataJuli2022/"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratN")){
    seuratN <- merge(x = seuratN, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Total cells:", ncol(seuratN), "\n")
  }else{
    seuratN <- seuratS
    cat("Initialized with first Seurat object:", fileNamList[i], "\n")
  }
}
remove(seuratS)
unique(seuratN$dataset)

##load objects June2019 dataNaive and merge all
basedir <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/dataJune2019/dataNaive/"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratN")){
    seuratN <- merge(x = seuratN, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Total cells:", ncol(seuratN), "\n")
  }else{
    seuratN <- seuratS
     cat("Initialized with first Seurat object:", fileNamList[i], "\n")
  }
}
remove(seuratS)
unique(seuratN$dataset)

##load objects Sept2020 and merge all
basedir <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/dataSept2020/"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratN")){
    seuratN <- merge(x = seuratN, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Total cells:", ncol(seuratN), "\n")
  }else{
    seuratN <- seuratS
  }
}
remove(seuratS)
unique(seuratN$dataset)

##load objects June2019 dataD8 and merge all
basedir <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/dataJune2019/dataD8/"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratN")){
    seuratN <- merge(x = seuratN, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Total cells:", ncol(seuratN), "\n")
  }else{
    seuratN <- seuratS
  }
}
remove(seuratS)
unique(seuratN$dataset)
table(seuratN$dataset)
table(seuratN$orig.ident)

##filer off adjacent and beads pool file
seuratN <- subset(seuratN, dataset %in% c("11_20200923_Mu_LP_Cxcl13Cre_IL33flfl_EYFP_HpD4__beads_pool_v3"), invert = T)
table(seuratN$dataset)
table(seuratN$orig.ident)

#rerun seurat
seuratN <- NormalizeData (object = seuratN)
seuratN <- FindVariableFeatures(object = seuratN)
seuratN <- ScaleData(object = seuratN, verbose = TRUE)
seuratN <- RunPCA(object=seuratN, npcs = 30, verbose = FALSE)
seuratN <- RunTSNE(object=seuratN, reduction="pca", dims = 1:20)
seuratN <- RunUMAP(object=seuratN, reduction="pca", dims = 1:20)
seuratN <- FindNeighbors(object = seuratN, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratN <- FindClusters(object = seuratN, resolution = res[i], random.seed = 1234)
}

saveRDS(seuratN, file="/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/merged_all_naiveANDd4ANDd8_seurat.rds")
```

## load object merged all
```{r load object merged all}
##load object merged all naive and d4 and d8
fileNam <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/merged_all_naiveANDd4ANDd8_seurat.rds"
seuratN <- readRDS(fileNam)
table(seuratN$dataset)
table(seuratN$orig.ident)
```

## plot umap all merged
```{r plot umap all merged}
## plot umap
Idents(seuratN) <- seuratN$RNA_snn_res.0.25
levels(seuratN)
DimPlot(seuratN, reduction = "umap", label = TRUE)

## plot feature
FeaturePlot(seuratN, features = "ENSMUSG00000026395.Ptprc", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratN, features = "ENSMUSG00000020717.Pecam1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratN, features = "ENSMUSG00000028583.Pdpn", pt.size = 1, cols = c("lightgrey", "#BE3144"))
```
## subset fibroblasts and rerun seurat
```{r subset fb and rerun seurat, eval=FALSE, include=TRUE}
## subset fb and rerun
seuratFb <- subset(seuratN, RNA_snn_res.0.25 %in% c("0", "1", "2", "5", "12", "11", "14", "4"))

## rerun seuratFb
seuratFb <- NormalizeData (object = seuratFb)
seuratFb <- FindVariableFeatures(object = seuratFb)
seuratFb <- ScaleData(object = seuratFb, verbose = TRUE)
seuratFb <- RunPCA(object=seuratFb, npcs = 30, verbose = FALSE)
seuratFb <- RunTSNE(object=seuratFb, reduction="pca", dims = 1:20)
seuratFb <- RunUMAP(object=seuratFb, reduction="pca", dims = 1:20)
seuratFb <- FindNeighbors(object = seuratFb, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratFb <- FindClusters(object = seuratFb, resolution = res[i], random.seed = 1234)
}

saveRDS(seuratFb, file="/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/merged_allFb_naiveANDd4ANDd8_seurat.rds")
```

## add metadata
```{r add metadata, eval=FALSE, include=TRUE}
## add any type of metadata
## cond2
WT <- c("o289411_1-1_20220629_Mu_granuloma_Cxcl13Cre_CD45_HpD4_v3",
        "o289411_2-2_20220629_Mu_granuloma_Cxcl13Cre_FSC_HpD4_v3",
        "9_20210624_Mu_granuloma_Cxcl13Cre_EYFP_HpD4_v3", 
        "11_20190604_Mouse_LP_Cxcl13cre_EYFP_v3", 
        "12_20190604_Mouse_LP_Cxcl13cre_EYFP_v3", 
        "9_20200923_Mu_LP_Cxcl13Cre_EYFP_HpD4_v3",
        "10_20200923_Mu_LP_Cxcl13Cre_EYFP_HpD4_v3",
        "15_20190613_Mouse_LP_Hp_d8_Cxcl13cre_EYFP_v3",
        "16_20190613_Mouse_LP_Hp_d8_Cxcl13cre_EYFP_v3")
cko <- c("o289411_3-3_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_CD45_HpD4_v3",
         "o289411_4-4_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_FSC_HpD4_v3",
         "10_20210624_Mu_granuloma_Cxcl13Cre_IL33flf_EYFP_HpD4_v3",
         "10_20190604_Mouse_LP_Cxcl13cre_IL33flfl_EYFP_v3",
         "9_20190604_Mouse_LP_Cxcl13cre_IL33flfl_EYFP_v3",
         "7_20200923_Mu_LP_Cxcl13Cre_IL33flfl_EYFP_HpD4_v3",
         "8_20200923_Mu_LP_Cxcl13Cre_IL33fllf_EYFP_HpD4_v3",
         "13_20190613_Mouse_LP_Hp_d8_Cxcl13cre_IL33flfl_EYFP",
          "14_20190613_Mouse_LP_Hp_d8_Cxcl13cre_IL33flfl_EYFP")
seuratFb$cond2 <- "cond2"
seuratFb$cond2[which(seuratFb$dataset %in% WT)] <- "WT"
seuratFb$cond2[which(seuratFb$dataset %in% cko)] <- "cko"
table(seuratFb$cond2)

## order
seuratFb$cond2 <- factor(seuratFb$cond2, levels=c("WT", "cko"))
Idents(seuratFb) <- seuratFb$cond2
table(seuratFb$cond2)

## mouse
m1wtnaive <- c("11_20190604_Mouse_LP_Cxcl13cre_EYFP_v3")
m2wtnaive <- c("12_20190604_Mouse_LP_Cxcl13cre_EYFP_v3")
m1ckonaive <- c("10_20190604_Mouse_LP_Cxcl13cre_IL33flfl_EYFP_v3")
m2ckonaive <- c("9_20190604_Mouse_LP_Cxcl13cre_IL33flfl_EYFP_v3")

m1wtD4SI <- c("9_20200923_Mu_LP_Cxcl13Cre_EYFP_HpD4_v3")
m2wtD4SI <- c("10_20200923_Mu_LP_Cxcl13Cre_EYFP_HpD4_v3")
m1ckoD4SI <- c("7_20200923_Mu_LP_Cxcl13Cre_IL33flfl_EYFP_HpD4_v3")
m2ckoD4SI <- c("8_20200923_Mu_LP_Cxcl13Cre_IL33fllf_EYFP_HpD4_v3")

m1wtD4Gran <- c("o289411_1-1_20220629_Mu_granuloma_Cxcl13Cre_CD45_HpD4_v3", "o289411_2-2_20220629_Mu_granuloma_Cxcl13Cre_FSC_HpD4_v3")
m2wtD4Gran <- c("9_20210624_Mu_granuloma_Cxcl13Cre_EYFP_HpD4_v3")
m1ckoD4Gran <- c("o289411_4-4_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_FSC_HpD4_v3","o289411_3-3_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_CD45_HpD4_v3")
m2ckoD4Gran <- c("10_20210624_Mu_granuloma_Cxcl13Cre_IL33flf_EYFP_HpD4_v3")

m1wtD8SI <- c("15_20190613_Mouse_LP_Hp_d8_Cxcl13cre_EYFP_v3")
m2wtD8SI <- c("16_20190613_Mouse_LP_Hp_d8_Cxcl13cre_EYFP_v3")
m1ckoD8SI <- c("13_20190613_Mouse_LP_Hp_d8_Cxcl13cre_IL33flfl_EYFP")
m2ckoD8SI <- c("14_20190613_Mouse_LP_Hp_d8_Cxcl13cre_IL33flfl_EYFP")

seuratFb$mouse <- "mouse"
seuratFb$mouse[which(seuratFb$dataset %in% m1wtnaive)] <- "m1wtnaive"
seuratFb$mouse[which(seuratFb$dataset %in% m2wtnaive)] <- "m2wtnaive"
seuratFb$mouse[which(seuratFb$dataset %in% m1ckonaive)] <- "m1ckonaive"
seuratFb$mouse[which(seuratFb$dataset %in% m2ckonaive)] <- "m2ckonaive"

seuratFb$mouse[which(seuratFb$dataset %in% m1wtD4SI)] <- "m1wtD4SI"
seuratFb$mouse[which(seuratFb$dataset %in% m2wtD4SI)] <- "m2wtD4SI"
seuratFb$mouse[which(seuratFb$dataset %in% m1ckoD4SI )] <- "m1ckoD4SI"
seuratFb$mouse[which(seuratFb$dataset %in% m2ckoD4SI)] <- "m2ckoD4SI"

seuratFb$mouse[which(seuratFb$dataset %in% m1wtD4Gran)] <- "m1wtD4Gran"
seuratFb$mouse[which(seuratFb$dataset %in% m2wtD4Gran)] <- "m2wtD4Gran"
seuratFb$mouse[which(seuratFb$dataset %in% m1ckoD4Gran)] <- "m1ckoD4Gran"
seuratFb$mouse[which(seuratFb$dataset %in% m2ckoD4Gran)] <- "m2ckoD4Gran"

seuratFb$mouse[which(seuratFb$dataset %in% m1wtD8SI)] <- "m1wtD8SI"
seuratFb$mouse[which(seuratFb$dataset %in% m2wtD8SI)] <- "m2wtD8SI"
seuratFb$mouse[which(seuratFb$dataset %in% m1ckoD8SI)] <- "m1ckoD8SI"
seuratFb$mouse[which(seuratFb$dataset %in% m2ckoD8SI)] <- "m2ckoD8SI"

## order
seuratFb$mouse <- factor(seuratFb$mouse, levels=c("m1wtnaive", "m2wtnaive", "m1ckonaive","m2ckonaive","m1wtD4SI", "m2wtD4SI", "m1ckoD4SI","m2ckoD4SI",  "m1wtD4Gran", "m2wtD4Gran",  "m1ckoD4Gran", "m2ckoD4Gran", "m1wtD8SI", "m2wtD8SI", "m1ckoD8SI", "m2ckoD8SI"))
Idents(seuratFb) <- seuratFb$mouse
table(seuratFb$mouse)

## Hp
d4 <- c("o289411_1-1_20220629_Mu_granuloma_Cxcl13Cre_CD45_HpD4_v3",
        "o289411_2-2_20220629_Mu_granuloma_Cxcl13Cre_FSC_HpD4_v3",
        "9_20210624_Mu_granuloma_Cxcl13Cre_EYFP_HpD4_v3",
        "o289411_3-3_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_CD45_HpD4_v3",
        "o289411_4-4_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_FSC_HpD4_v3",
        "10_20210624_Mu_granuloma_Cxcl13Cre_IL33flf_EYFP_HpD4_v3",
        "9_20200923_Mu_LP_Cxcl13Cre_EYFP_HpD4_v3",
        "10_20200923_Mu_LP_Cxcl13Cre_EYFP_HpD4_v3",
        "7_20200923_Mu_LP_Cxcl13Cre_IL33flfl_EYFP_HpD4_v3",
        "8_20200923_Mu_LP_Cxcl13Cre_IL33fllf_EYFP_HpD4_v3")
d8 <- c("15_20190613_Mouse_LP_Hp_d8_Cxcl13cre_EYFP_v3",
        "16_20190613_Mouse_LP_Hp_d8_Cxcl13cre_EYFP_v3",
        "13_20190613_Mouse_LP_Hp_d8_Cxcl13cre_IL33flfl_EYFP",
          "14_20190613_Mouse_LP_Hp_d8_Cxcl13cre_IL33flfl_EYFP")
naive <- c("11_20190604_Mouse_LP_Cxcl13cre_EYFP_v3", 
        "12_20190604_Mouse_LP_Cxcl13cre_EYFP_v3",
        "10_20190604_Mouse_LP_Cxcl13cre_IL33flfl_EYFP_v3",
        "9_20190604_Mouse_LP_Cxcl13cre_IL33flfl_EYFP_v3")
seuratFb$Hp <- "Hp"
seuratFb$Hp[which(seuratFb$dataset %in% d4)] <- "d4"
seuratFb$Hp[which(seuratFb$dataset %in% d8)] <- "d8"
seuratFb$Hp[which(seuratFb$dataset %in% naive)] <- "naive"
table(seuratFb$Hp)

## cond3
infectedD4 <- c("9_20200923_Mu_LP_Cxcl13Cre_EYFP_HpD4_v3",
        "10_20200923_Mu_LP_Cxcl13Cre_EYFP_HpD4_v3",
        "7_20200923_Mu_LP_Cxcl13Cre_IL33flfl_EYFP_HpD4_v3",
        "8_20200923_Mu_LP_Cxcl13Cre_IL33fllf_EYFP_HpD4_v3")
infectedD8 <- c("15_20190613_Mouse_LP_Hp_d8_Cxcl13cre_EYFP_v3",
        "16_20190613_Mouse_LP_Hp_d8_Cxcl13cre_EYFP_v3",
        "13_20190613_Mouse_LP_Hp_d8_Cxcl13cre_IL33flfl_EYFP",
          "14_20190613_Mouse_LP_Hp_d8_Cxcl13cre_IL33flfl_EYFP")
granulomaD4 <- c("o289411_1-1_20220629_Mu_granuloma_Cxcl13Cre_CD45_HpD4_v3",
        "o289411_2-2_20220629_Mu_granuloma_Cxcl13Cre_FSC_HpD4_v3",
        "9_20210624_Mu_granuloma_Cxcl13Cre_EYFP_HpD4_v3",
        "o289411_3-3_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_CD45_HpD4_v3",
        "o289411_4-4_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_FSC_HpD4_v3",
        "10_20210624_Mu_granuloma_Cxcl13Cre_IL33flf_EYFP_HpD4_v3")
naive <- c("11_20190604_Mouse_LP_Cxcl13cre_EYFP_v3", 
        "12_20190604_Mouse_LP_Cxcl13cre_EYFP_v3",
        "10_20190604_Mouse_LP_Cxcl13cre_IL33flfl_EYFP_v3",
        "9_20190604_Mouse_LP_Cxcl13cre_IL33flfl_EYFP_v3")
seuratFb$cond3 <- "cond3"
seuratFb$cond3[which(seuratFb$dataset %in% infectedD4)] <- "infectedD4"
seuratFb$cond3[which(seuratFb$dataset %in% infectedD8)] <- "infectedD8"
seuratFb$cond3[which(seuratFb$dataset %in% granulomaD4)] <- "granulomaD4"
seuratFb$cond3[which(seuratFb$dataset %in% naive)] <- "naive"
table(seuratFb$cond3)

## order
seuratFb$cond3 <- factor(seuratFb$cond3, levels=c("naive", "infectedD4", "granulomaD4", "infectedD8"))
Idents(seuratFb) <- seuratFb$cond3
table(seuratFb$cond3)

seuratFb$cond2_plus_cond3 <- paste0(seuratFb$cond2, "_", seuratFb$cond3)
table(seuratFb$cond2_plus_cond3)

## clustername
seuratFb$clustername <- "clustername"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "0" )] <- "PdgfraloFb1"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "1" )] <- "PdgfraloFb2"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "2" )] <- "Trophocytes"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "3" )] <- "Telocytes"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "4" )] <- "PdgfraloFb3"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "5" )] <- "PdgfraloFb4"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "6" )] <- "Myocytes"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "7" )] <- "Pdgfrahi"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "8" )] <- "Prolif"
seuratFb$clustername[which(seuratFb$RNA_snn_res.0.25 %in% "9" )] <- "Thy1Fb"

table(seuratFb$clustername)
table(seuratFb$RNA_snn_res.0.25)

## order
seuratFb$clustername <- factor(seuratFb$clustername, levels=c("PdgfraloFb1","PdgfraloFb2","PdgfraloFb3","PdgfraloFb4","Trophocytes","Telocytes","Myocytes", "Pdgfrahi", "Thy1Fb", "Prolif"))
Idents(seuratFb) <- seuratFb$clustername
table(seuratFb$clustername)

seuratFb$clustername_plus_cond3 <- paste0(seuratFb$clustername, "_", seuratFb$cond3)
table(seuratFb$clustername_plus_cond3)

### save seurat object
saveRDS(seuratFb, file="/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/merged_allFb_naiveANDd4ANDd8_seurat.rds")
```
#################################################

# explore Fb

## load object merged allFb 
```{r load object merged allFb}
##load object merged allFb naive and d4 and d8
fileNam <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/merged_allFb_naiveANDd4ANDd8_seurat.rds"
seuratFb <- readRDS(fileNam)
table(seuratFb$dataset)
```

## set color vectors 
```{r colors}
## set vector for cond2
colcond2 <- c("#202547","#BE3144")
names(colcond2) <- c("WT", "cko")

## set vector for cond2
colcond3 <- c("#B45B5C","#628395","#BF782D","#67001f")# "#f4a582"
names(colcond3) <- c("infectedD4", "naive", "granulomaD4", "infectedD8")

#colPal <- c("#355C7D", "#8E9B97","#779d8d","#0F1F38","#E84A5F","#FF847C","#F8B195", "#727077", "#C06C84","#2A363B", "#6C5B7B")
#names(colPal) <- c("0" ,"1", "2", "3", "4", "5", "6", "7", "8", "9", "10")

colPal <- c("#355C7D", "#8E9B97","#99B898","#E84A5F","#0F1F38","#FF847C","#F8B195","#6C5B7B", "#C06C84", "#727077")
names(colPal) <- c("0" ,"1", "2", "3","4", "5", "6", "7", "8", "9")

colclustername <- c("#355C7D", "#8E9B97","#99B898","#E84A5F","#0F1F38","#F8B195","#727077","#FF847C","#C06C84","#904D39")
names(colclustername) <- c("PdgfraloFb1" ,"PdgfraloFb2", "Trophocytes","PdgfraloFb3", "Telocytes","Myocytes", "Pdgfrahi", "PdgfraloFb4", "Thy1Fb", "Prolif")
```

## umaps clustername
```{r umaps clustername}
## RNA_snn_res.0.25
Idents(seuratFb) <- seuratFb$RNA_snn_res.0.25
DimPlot(seuratFb, reduction= "umap", label = TRUE)
## clustername
Idents(seuratFb) <- seuratFb$clustername
DimPlot(seuratFb, reduction= "umap", cols = colclustername)
DimPlot(seuratFb, reduction= "umap", cols = colclustername) + theme(legend.position = "none")
## cond2
Idents(seuratFb) <- seuratFb$cond2
DimPlot(seuratFb, reduction= "umap", cols = colcond2, shuffle = TRUE)
## cond3
Idents(seuratFb) <- seuratFb$cond3
DimPlot(seuratFb, reduction= "umap", cols = colcond3, shuffle = TRUE)
DimPlot(seuratFb, reduction= "umap", cols = colcond3, shuffle = TRUE, split.by = "cond3")
```

## calculate cluster marker genes
```{r marker gens, include=TRUE, eval=FALSE}
## calculate marker genes
Idents(seuratFb) <- seuratFb$clustername
levels(seuratFb)

markerGenes <- FindAllMarkers(seuratFb, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01) 

write.table(markerGenes, 
            file= "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/FbmarkerGenesclusternameincl.D8.txt",
            sep="\t",
            quote=F,
            row.names=F,
            col.names=T)
```

## dotplot fb
```{r dotplot fb, fig.height=4, fig.width=7}
genes <- data.frame(gene=rownames(seuratFb)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

selGenes <- data.frame(geneID=c("Thy1","Ly6a","Acta2","Cd81","Cd34","Pdgfra","Pdpn")) %>% left_join(., genes, by="geneID")

DotPlot(seuratFb, features = selGenes, group.by= "clustername") + RotatedAxis() + scale_color_viridis(option="E") + coord_flip()
DotPlot(seuratFb, features = selGenes, group.by= "clustername") + RotatedAxis() + scale_color_viridis(option="F") + coord_flip()
DotPlot(seuratFb, features = selGenes, group.by= "clustername") + RotatedAxis() + scale_color_viridis(option="D") + coord_flip()

selGenes <- data.frame(geneID=c("Mki67","Ncam1","Ly6a","Acta2","Cd81","Cd34","Pdgfra","Pdpn")) %>% left_join(., genes, by="geneID")

DotPlot(seuratFb, features = selGenes, group.by= "clustername") + RotatedAxis() + scale_color_viridis(option="E") + coord_flip()
DotPlot(seuratFb, features = selGenes, group.by= "clustername") + RotatedAxis() + scale_color_viridis(option="F") + coord_flip()
DotPlot(seuratFb, features = selGenes, group.by= "clustername") + RotatedAxis() + scale_color_viridis(option="D") + coord_flip()
```

## abundances Fb subsets
```{r Fb abundances plot}
## make count list according to cond3
datList <- NULL
for(con in unique(seuratFb$cond3)){
  seuratSub <- subset(seuratFb, cond3==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clustername)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

##order x
ordX <-  c("naive", "infectedD4", "granulomaD4", "infectedD8") 

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "clustername", ylab = "frequency", palette = colclustername) + scale_x_discrete(limits=ordX)

## make count list according to cond2 plus cond3
datList <- NULL
for(con in unique(seuratFb$cond2_plus_cond3)){
  seuratSub <- subset(seuratFb, cond2_plus_cond3==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clustername)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

##order x
ordX <-  c("WT_naive", "cko_naive", "WT_infectedD4", "cko_infectedD4", "WT_granulomaD4", "cko_granulomaD4", "WT_infectedD8", "cko_infectedD8") 

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "clustername", ylab = "frequency", palette = colclustername) + scale_x_discrete(limits=ordX) + theme(axis.text.x = element_text(angle = 20, hjust=1))

## make count list according to mouse
datList <- NULL
for(con in unique(seuratFb$mouse)){
  seuratSub <- subset(seuratFb, mouse==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clustername)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

##order x
ordX <- c("m1wtnaive", "m2wtnaive", "m1ckonaive","m2ckonaive","m1wtD4SI", "m2wtD4SI", "m1ckoD4SI","m2ckoD4SI",  "m1wtD4Gran", "m2wtD4Gran",  "m1ckoD4Gran", "m2ckoD4Gran", "m1ckoD8SI","m2ckoD8SI", "m1wtD8SI","m2wtD8SI")

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "clustername", ylab = "frequency", palette = colclustername) + theme(axis.text.x = element_text(angle = 20, hjust=1)) + scale_x_discrete(limits=ordX) 
```

### subset fibroblast wt
```{r subset seurat Fb wt, fig.height=8, fig.width=15}
table(seuratFb$cond2)
seuratFbwt <- subset(seuratFb, cond2 == "WT")
table(seuratFbwt$orig.ident)
```

## avg heatmap cluster marker genes (Prados, Onder, Nat Immunol. 2021)
```{r avg heatmap PP FRC cluster marker}
Idents(seuratFb) <- seuratFb$cond3
levels(seuratFb)
seurat <- seuratFb
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     #ordVec <- c("naive","infectedD4","granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
#	cytokine-mediated signaling pathway
#negative regulation of endopeptidase activity
genesPlot <- c("Cttnbp2","Cr2", "Cxcl13", "Madcam1", "Acta2", "Cxcl12", "Il33", "Il7", "Ccl19", "Ccl21a", "Col1a1", "C3", "Ltbr", "Tnfrsf1a", "Irf7", "Ifit3", "Ifit1")

levels(seurat)
colVec <- colcond3
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

## avg heatmap cluster marker genes (Prados, Onder, Nat Immunol. 2021)
```{r avg heatmap PP FRC cluster marker-2}
Idents(seuratFb) <- seuratFb$cond2_plus_cond3
levels(seuratFb)
seurat <- seuratFb
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     ordVec <- c("WT_naive","cko_naive", "WT_infectedD4","cko_infectedD4","WT_granulomaD4", "cko_granulomaD4", "WT_infectedD8", "cko_infectedD8")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
#	cytokine-mediated signaling pathway
#negative regulation of endopeptidase activity
genesPlot <- c("Cttnbp2","Cr2", "Cxcl13", "Madcam1", "Acta2", "Cxcl12", "Il33", "Il7", "Ccl19", "Ccl21a", "Col1a1", "C3", "Ltbr", "Tnfrsf1a", "Irf7", "Ifit3", "Ifit1")

levels(seurat)
colVec <- colcond2
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

## avg heatmap cluster marker genes (Cheng et al. Nat. commun. 2022)
```{r avg heatmap SILT FRC cluster marker}
Idents(seuratFb) <- seuratFb$cond3
levels(seuratFb)
seurat <- seuratFb
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     #ordVec <- c("naive","infectedD4","granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

#genes
genesPlot <- c("Clu", "Il7", "Vcam1", "Icam1", "Cxcl16", "Cxcl13", "Cxcl10", "Cxcl9", "Ccl2", "Tnfsf13b", "Tnfsf11", "Kitl")

levels(seurat)
colVec <- colcond3
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

## avg heatmap cluster marker genes (Cheng et al. Nat. commun. 2022)
```{r avg heatmap SILT FRC cluster marker-2}
Idents(seuratFb) <- seuratFb$cond2_plus_cond3
levels(seuratFb)
seurat <- seuratFb
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     ordVec <- c("WT_naive","cko_naive", "WT_infectedD4","cko_infectedD4","WT_granulomaD4", "cko_granulomaD4", "WT_infectedD8", "cko_infectedD8")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
genesPlot <- c("Clu", "Il7", "Vcam1", "Icam1", "Cxcl16", "Cxcl13", "Cxcl10", "Cxcl9", "Ccl2", "Tnfsf13b", "Tnfsf11", "Kitl")

levels(seurat)
colVec <- colcond2
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

## violin plots of FRC genes
```{r violin plots of FRC genes}
Idents(seuratFbwt) <- seuratFbwt$cond3
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000071005.Ccl19", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000094686.Ccl21a", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000035385.Ccl2", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000040329.Il7", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000022015.Tnfsf11", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000020310.Madcam1", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000022037.Clu", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000019966.Kitl", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000030339.Ltbr", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000025498.Irf7", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000031497.Tnfsf13b", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000029417.Cxcl9", pt.size = 0, cols = colcond3) 
```

## session info
```{r date and session info}
date()
sessionInfo()
```
