---
title: "Fibroblasts wt only"
author: "A.DeMartin"
date: "2024-02-21"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages
```{r load packages}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Mm.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
```

############################ end pre-processing############################

## load object merged allFb d4 
```{r load object merged allFb d4}
##load object merged allFb naive and d4
fileNam <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/merged_allFb_naiveANDd4_seurat.rds"
seuratFb <- readRDS(fileNam)
table(seuratFb$dataset)
```

## subset wt
```{r subset wt}
table(seuratFb$cond2)
seuratwt <- subset(seuratFb, cond2 == "WT")
table(seuratwt$cond2)

seuratFb <- seuratwt
```

## set color vectors 
```{r colors}
## set vector for cond2
colcond2 <- c("#202547","#BE3144")
names(colcond2) <- c("WT", "cko")

## set vector for cond2
colcond3 <- c("#B45B5C","#628395","#BF782D")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")

#colPal <- c("#355C7D", "#8E9B97","#779d8d","#0F1F38","#E84A5F","#FF847C","#F8B195", "#727077", "#C06C84","#2A363B", "#6C5B7B")
#names(colPal) <- c("0" ,"1", "2", "3", "4", "5", "6", "7", "8", "9", "10")

colPal <- c("#355C7D", "#8E9B97","#99B898","#E84A5F","#0F1F38","#FF847C","#F8B195","#6C5B7B", "#C06C84", "#727077")
names(colPal) <- c("0" ,"1", "2", "3","4", "5", "6", "7", "8", "9")

colclustername <- c("#355C7D", "#8E9B97","#99B898","#E84A5F","#0F1F38","#F8B195","#727077","#FF847C","#C06C84","#904D39")
names(colclustername) <- c("PdgfraloFb1" ,"PdgfraloFb2", "Trophocytes","PdgfraloFb3", "Telocytes","Myocytes", "Pdgfrahi", "PdgfraloFb4", "Thy1Fb", "Glial")
```

## umaps clustername
```{r umaps clustername}
## cluster name
Idents(seuratFb) <- seuratFb$clustername
table(seuratFb$clustername)
table(seuratFb$orig.ident)
table(seuratFb$cond3)
DimPlot(seuratFb, reduction= "umap", cols = colclustername)
DimPlot(seuratFb, reduction= "umap", cols = colclustername) + theme(legend.position = "none")
```


## umaps conditions
```{r umaps conditions}
## cond3
Idents(seuratFb) <- seuratFb$cond3
table(seuratFb$cond3)
DimPlot(seuratFb, reduction= "umap", cols = colcond3, shuffle = T)
DimPlot(seuratFb, reduction= "umap", cols = colcond3, shuffle = T) + theme(legend.position = "none")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, split.by = "cond3") + theme(legend.position = "none")

## on grey background 
colcond3 <- c("lightgrey","lightgrey","#BF782D")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "granulomaD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "granulomaD4") + theme(legend.position = "none")

colcond3 <- c("lightgrey","#628395","lightgrey")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "naive")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "naive") + theme(legend.position = "none")

colcond3 <- c("#B45B5C","lightgrey","lightgrey")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "infectedD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "infectedD4") + theme(legend.position = "none")

colcond3 <- c("#B45B5C","#628395","#BF782D")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")
seuratFbgran <- subset(seuratFb, cond3 == "granulomaD4")
DimPlot(seuratFbgran, reduction= "umap", cols = colcond3) + theme(legend.position = "none")

seuratFbinf <- subset(seuratFb, cond3 == "infectedD4")
DimPlot(seuratFbinf, reduction= "umap", cols = colcond3) + theme(legend.position = "none")

seuratFbnaive <- subset(seuratFb, cond3 == "naive")
DimPlot(seuratFbnaive, reduction= "umap", cols = colcond3) + theme(legend.position = "none")

## cond2
Idents(seuratFb) <- seuratFb$cond2
table(seuratFb$cond2)
DimPlot(seuratFb, reduction= "umap", cols = colcond2)
DimPlot(seuratFb, reduction= "umap", cols = colcond2, shuffle = T)
DimPlot(seuratFb, reduction= "umap", cols = colcond2, split.by = "cond2") + theme(legend.position = "none")
```

## feature plots fb maker genes
```{r feature plots fb}
## plot feature
FeaturePlot(seuratFb, features = "ENSMUSG00000028583.Pdpn", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000028583.Pdpn", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "Rosa26eyfp.Rosa26eyfp", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "Rosa26eyfp.Rosa26eyfp", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000035783.Acta2", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000035783.Acta2", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000016494.Cd34", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000016494.Cd34", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000029231.Pdgfra", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000029231.Pdgfra", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000023078.Cxcl13", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000023078.Cxcl13", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000071005.Ccl19", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000071005.Ccl19", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000039542.Ncam1", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000039542.Ncam1", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000075602.Ly6a", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000075602.Ly6a", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000037706.Cd81", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000037706.Cd81", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000020717.Pecam1", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000020717.Pecam1", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000045394.Epcam", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000045394.Epcam", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000024810.Il33", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000024810.Il33", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000023078.Cxcl13", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000023078.Cxcl13", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000022057.Adamdec1", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000022057.Adamdec1", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000020676.Ccl11", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000020676.Ccl11", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000024011.Pi16", cols = c("lightgrey", "#BE3144"), pt.size = 1, order = TRUE)
FeaturePlot(seuratFb, features = "ENSMUSG00000024011.Pi16", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000094686.Ccl21a", cols = c("lightgrey", "#BE3144"), pt.size = 1, order = TRUE)
FeaturePlot(seuratFb, features = "ENSMUSG00000094686.Ccl21a", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000027358.Bmp2", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000027358.Bmp2", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)
```

## average Heatmap - Fb top5 cluster marker
```{r avg heatmap, fig.height=10, fig.width=8}
##order
Idents(seuratFb) <- seuratFb$clustername
levels(seuratFb)

seurat <- seuratFb
avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colclustername)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

genesPlot <- data.frame(gene=c("Rgs5", "Apcdd1", "Timp3", "Cd34", "Shisa3", "Igfbp3", "Mmp10", "Agt", "Lpl", "Fhl2", "Cxcl5", "Chl1", "Serpina3g", "Ecscr", "Slc15a3", "Itga8", "P2rx1", "Tmem178", "Plvap", "Il34", "Cd81", "Ackr4", "Ogn", "Pcolce2", "Scara5", "Sox6", "Kcnj8", "Tmem158", "Tspan13", "Bmp7", "Actg2", "Myh11", "Lmod1", "Npnt", "Tagln", "Ptn", "Tmem229a", "Fxyd6", "Chchd10", "Cox8b", "Pla1a", "Ace", "Ifi207", "Plau", "Prrx1", "Fabp5", "Plp1", "Pdgfb", "S100a4", "Sostdc1"))

levels(seurat)
colVec <- colPal
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

### dotplot fb
```{r dotplot fb, fig.height=4, fig.width=7}
genes <- data.frame(gene=rownames(seuratFb)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

selGenes <- data.frame(geneID=c("Il33", "Thy1","Ly6a","Acta2","Cd81","Cd34","Pdgfra","Pdpn")) %>% left_join(., genes, by="geneID")

DotPlot(seuratFb, features = selGenes, group.by= "cluster_name") + RotatedAxis() + scale_color_viridis(option="E") + coord_flip()
DotPlot(seuratFb, features = selGenes, group.by= "cluster_name") + RotatedAxis() + scale_color_viridis(option="F") + coord_flip()
```

## Rel Abundance Fb subsets
```{r Fb abundance plot}
## make count list according to cond3
datList <- NULL
for(con in unique(seuratFb$cond3)){
  seuratSub <- subset(seuratFb, cond3==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clustername)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

##order x
ordX <-  c("naive", "infectedD4", "granulomaD4") 

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "clustername", ylab = "frequency", palette = colclustername) + scale_x_discrete(limits=ordX)

## make count list according to mouse
datList <- NULL
for(con in unique(seuratFb$mouse)){
  seuratSub <- subset(seuratFb, mouse==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clustername)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

##order x
ordX <- c("m1wtnaive", "m2wtnaive","m1wtD4SI", "m2wtD4SI",  "m1wtD4Gran", "m2wtD4Gran")

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "clustername", ylab = "frequency", palette = colclustername) + theme(axis.text.x = element_text(angle = 20, hjust=1)) + scale_x_discrete(limits=ordX) 
```

## without gial cells
```{r}
##filter glial
seuratFb_noglial <- subset(seuratFb, clustername == "Glial", invert = TRUE)

table(seuratFb$clustername)
table(seuratFb_noglial$clustername)

table(seuratFb$orig.ident)
table(seuratFb_noglial$orig.ident)

seuratFb <- seuratFb_noglial
```

## umaps clustername
```{r umaps clustername no glial}
## cluster name
Idents(seuratFb) <- seuratFb$clustername
table(seuratFb$clustername)
table(seuratFb$orig.ident)
table(seuratFb$cond3)
DimPlot(seuratFb, reduction= "umap", cols = colclustername)
DimPlot(seuratFb, reduction= "umap", cols = colclustername) + theme(legend.position = "none")
```

## umaps conditions
```{r umaps conditions no glial}
## cond3
Idents(seuratFb) <- seuratFb$cond3
table(seuratFb$cond3)
DimPlot(seuratFb, reduction= "umap", cols = colcond3, shuffle = T)
DimPlot(seuratFb, reduction= "umap", cols = colcond3, shuffle = T) + theme(legend.position = "none")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, split.by = "cond3") + theme(legend.position = "none")

## on grey background 
colcond3 <- c("lightgrey","lightgrey","#BF782D")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "granulomaD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "granulomaD4") + theme(legend.position = "none")

colcond3 <- c("lightgrey","#628395","lightgrey")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "naive")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "naive") + theme(legend.position = "none")

colcond3 <- c("#B45B5C","lightgrey","lightgrey")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "infectedD4")
DimPlot(seuratFb, reduction= "umap", cols = colcond3, order = "infectedD4") + theme(legend.position = "none")

colcond3 <- c("#B45B5C","#628395","#BF782D")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")
seuratFbgran <- subset(seuratFb, cond3 == "granulomaD4")
DimPlot(seuratFbgran, reduction= "umap", cols = colcond3) + theme(legend.position = "none")

seuratFbinf <- subset(seuratFb, cond3 == "infectedD4")
DimPlot(seuratFbinf, reduction= "umap", cols = colcond3) + theme(legend.position = "none")

seuratFbnaive <- subset(seuratFb, cond3 == "naive")
DimPlot(seuratFbnaive, reduction= "umap", cols = colcond3) + theme(legend.position = "none")

## cond2
Idents(seuratFb) <- seuratFb$cond2
table(seuratFb$cond2)
DimPlot(seuratFb, reduction= "umap", cols = colcond2)
DimPlot(seuratFb, reduction= "umap", cols = colcond2, shuffle = T)
DimPlot(seuratFb, reduction= "umap", cols = colcond2, split.by = "cond2") + theme(legend.position = "none")
```

## feature plots fb maker genes
```{r feature plots fb no glial}
## plot feature
FeaturePlot(seuratFb, features = "ENSMUSG00000028583.Pdpn", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000028583.Pdpn", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "Rosa26eyfp.Rosa26eyfp", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "Rosa26eyfp.Rosa26eyfp", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000035783.Acta2", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000035783.Acta2", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000016494.Cd34", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000016494.Cd34", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000029231.Pdgfra", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000029231.Pdgfra", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000023078.Cxcl13", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000023078.Cxcl13", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000071005.Ccl19", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000071005.Ccl19", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000039542.Ncam1", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000039542.Ncam1", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000075602.Ly6a", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000075602.Ly6a", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000037706.Cd81", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000037706.Cd81", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000020717.Pecam1", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000020717.Pecam1", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000045394.Epcam", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000045394.Epcam", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000024810.Il33", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000024810.Il33", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000023078.Cxcl13", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000023078.Cxcl13", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000022057.Adamdec1", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000022057.Adamdec1", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000020676.Ccl11", cols = c("lightgrey", "#BE3144"), pt.size = 1)
FeaturePlot(seuratFb, features = "ENSMUSG00000020676.Ccl11", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000024011.Pi16", cols = c("lightgrey", "#BE3144"), pt.size = 1, order = TRUE)
FeaturePlot(seuratFb, features = "ENSMUSG00000024011.Pi16", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)

FeaturePlot(seuratFb, features = "ENSMUSG00000094686.Ccl21a", cols = c("lightgrey", "#BE3144"), pt.size = 1, order = TRUE)
FeaturePlot(seuratFb, features = "ENSMUSG00000094686.Ccl21a", cols = c("lightgrey", "#BE3144"), pt.size = 1) + theme(legend.position = "none") + ggtitle(NULL)
```

## average Heatmap - Fb top5 cluster marker
```{r avg heatmap no glial, fig.height=10, fig.width=8}
##order
Idents(seuratFb) <- seuratFb$clustername
levels(seuratFb)

seurat <- seuratFb
avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colclustername)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

genesPlot <- data.frame(gene=c("Rgs5", "Apcdd1", "Timp3", "Cd34", "Shisa3", "Igfbp3", "Mmp10", "Agt", "Lpl", "Fhl2", "Cxcl5", "Chl1", "Serpina3g", "Ecscr", "Slc15a3", "Itga8", "P2rx1", "Tmem178", "Plvap", "Il34", "Cd81", "Ackr4", "Ogn", "Pcolce2", "Scara5", "Sox6", "Kcnj8", "Tmem158", "Tspan13", "Bmp7", "Actg2", "Myh11", "Lmod1", "Npnt", "Tagln", "Ptn", "Tmem229a", "Fxyd6", "Chchd10", "Cox8b", "Pla1a", "Ace", "Ifi207", "Plau", "Prrx1", "Fabp5", "Plp1", "Pdgfb", "S100a4", "Sostdc1"))

levels(seurat)
colVec <- colPal
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

### dotplot fb
```{r dotplot fb no glial, fig.height=4, fig.width=7}
genes <- data.frame(gene=rownames(seuratFb)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

selGenes <- data.frame(geneID=c("Il33","Thy1","Ly6a","Acta2","Cd81","Cd34","Pdgfra","Pdpn")) %>% left_join(., genes, by="geneID")

DotPlot(seuratFb, features = selGenes, group.by= "cluster_name") + RotatedAxis() + scale_color_viridis(option="E") + coord_flip()
DotPlot(seuratFb, features = selGenes, group.by= "cluster_name") + RotatedAxis() + scale_color_viridis(option="F") + coord_flip()
```

## Rel Abundance Fb subsets
```{r Fb abundance plot no glial}
## make count list according to cond3
datList <- NULL
for(con in unique(seuratFb$cond3)){
  seuratSub <- subset(seuratFb, cond3==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clustername)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

##order x
ordX <-  c("naive", "infectedD4", "granulomaD4") 

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "clustername", ylab = "frequency", palette = colclustername) + scale_x_discrete(limits=ordX)

## make count list according to mouse
datList <- NULL
for(con in unique(seuratFb$mouse)){
  seuratSub <- subset(seuratFb, mouse==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clustername)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

##order x
ordX <- c("m1wtnaive", "m2wtnaive","m1wtD4SI", "m2wtD4SI",  "m1wtD4Gran", "m2wtD4Gran")

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "clustername", ylab = "frequency", palette = colclustername) + theme(axis.text.x = element_text(angle = 20, hjust=1)) + scale_x_discrete(limits=ordX) 
```

## session info
```{r date and session info}
date()
sessionInfo()
```
