---
title: "FRCsignatures"
author: "A.DeMartin"
date: "2025-04-01"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages
```{r load packages}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Mm.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
```

## load object merged allFb d4
```{r load object merged allFb d4}
##load object merged allFb naive and d4
fileNam <- "/Users/immbio/Desktop/Project/Angelina/Cxcl13Hpb/data/merged_allFb_naiveANDd4_seurat.rds"
seuratFb <- readRDS(fileNam)
table(seuratFb$dataset)
```

### subset fibroblast wt
```{r subset seurat Fb wt, fig.height=8, fig.width=15}
table(seuratFb$cond2)
seuratFbwt <- subset(seuratFb, cond2 == "WT")
table(seuratFbwt$orig.ident)

table(seuratFbwt$clustername)
seuratnoglial <- subset(seuratFbwt, clustername == "Glial", invert = TRUE)
table(seuratnoglial$clustername)

seuratFbwt <- seuratnoglial
```

## set color vectors 
```{r colors}
## set vector for cond2
colcond2 <- c("#202547","#BE3144")
names(colcond2) <- c("WT", "cko")

## set vector for cond2
colcond3 <- c("#B45B5C","#628395","#BF782D")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")

#colPal <- c("#355C7D", "#8E9B97","#779d8d","#0F1F38","#E84A5F","#FF847C","#F8B195", "#727077", "#C06C84","#2A363B", "#6C5B7B")
#names(colPal) <- c("0" ,"1", "2", "3", "4", "5", "6", "7", "8", "9", "10")

colPal <- c("#355C7D", "#8E9B97","#99B898","#E84A5F","#0F1F38","#FF847C","#F8B195","#6C5B7B", "#C06C84", "#727077")
names(colPal) <- c("0" ,"1", "2", "3","4", "5", "6", "7", "8", "9")

colclustername <- c("#355C7D", "#8E9B97","#99B898","#E84A5F","#0F1F38","#F8B195","#727077","#FF847C","#C06C84","#904D39")
names(colclustername) <- c("PdgfraloFb1" ,"PdgfraloFb2", "Trophocytes","PdgfraloFb3", "Telocytes","Myocytes", "Pdgfrahi", "PdgfraloFb4", "Thy1Fb", "Glial")
```

### heatmap of cluster marker genes (Prados, Onder, Nat Immunol. 2021)
```{r project PP FRC gene signatures, fig.height=8, fig.width=15}
Idents(seuratFbwt) <- seuratFbwt$clustername
genes <- data.frame(gene=rownames(seuratFb)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

PPclgenes <- data.frame(geneID=(c("Igfbp3", "Des", "Mme", "Il33", "Ccl19", "Plvap", "Aqp1", "Ccl21a", "Mustn1", "Enpp2", "Acta2", "Tagln", "Igfbp5", "Vtn", "Cd74", "Cst3", "Thbs2", "Lum", "Ccn5", "C3", "Serpine2", "Igfbp4", "Fmod", "Ccl7", "Cxcl5", "Ifit3", "Pcolce2", "Cd34", "Fmo2", "Clec3b", "Fbln1", "Ackr4", "Gsn", "Col14a1", "Col3a1", "Col1a1", "Col5a1", "Ly6a", "Klf4", "Egr1", "Mt1", "Sox9", "Brinp2", "Slc1a5", "Slc1a2", "Slc41a1", "Relt", "Mef2c", "Cttnbp2", "Malat1", "Pde4d", "Fcamr", "Epcam", "Aqp4", "Csn2", "Coch","Serpina1e", "Il4i1", "Tmem119", "Sncg", "Fabp5", "Ubd", "Srgn", "Ctsd", "Cstb","Atox1"))) %>% left_join(., genes, by="geneID")

DoHeatmap(seuratFbwt, features = PPclgenes$gene, group.colors = colclustername) +
  scale_fill_gradientn(colors = c("#2166AC", "#F7F7F7", "#B2182B"))

Idents(seuratFbwt) <- seuratFbwt$cond3
DoHeatmap(seuratFbwt, features = PPclgenes$gene, group.colors = colcond3) +
  scale_fill_gradientn(colors = c("#2166AC", "#F7F7F7", "#B2182B"))
```

## avg heatmap cluster marker genes (Prados, Onder, Nat Immunol. 2021)
```{r avg heatmap PP FRC cluster marker}
Idents(seuratFbwt) <- seuratFbwt$cond3
levels(seuratFbwt)
seurat <- seuratFbwt
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     #ordVec <- c("naive","infectedD4","granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
#	cytokine-mediated signaling pathway
#negative regulation of endopeptidase activity
genesPlot <- c("Cttnbp2","Cr2", "Cxcl13", "Madcam1", "Acta2", "Cxcl12", "Il33", "Il7", "Ccl19", "Ccl21a", "Col1a1", "C3", "Ltbr", "Tnfrsf1a", "Irf7", "Ifit3", "Ifit1")

levels(seurat)
colVec <- colcond3
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

### avg heatmap cluster marker genes (Prados, Onder, Nat Immunol. 2021)
```{r avg heatmap PP FRC cluster marker-2}
Idents(seuratFbwt) <- seuratFbwt$clustername
levels(seuratFbwt)
seurat <- seuratFbwt
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     #ordVec <- c("naive","infectedD4","granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
genesPlot <- c("Cttnbp2","Cr2", "Cxcl13", "Madcam1", "Acta2", "Cxcl12", "Il33", "Il7", "Ccl19", "Ccl21a", "Col1a1", "C3", "Ltbr", "Tnfrsf1a", "Irf7", "Ifit3", "Ifit1")

levels(seurat)
colVec <- colclustername
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

### avg heatmap cluster marker genes (Cheng et al. Nat. commun. 2022)
```{r avg heatmap SILT FRC cluster marker}
Idents(seuratFbwt) <- seuratFbwt$cond3
levels(seuratFbwt)
seurat <- seuratFbwt
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     #ordVec <- c("naive","infectedD4","granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
genesPlot <- c("Clu", "Il7", "Vcam1", "Icam1", "Cxcl16", "Cxcl13", "Cxcl10", "Cxcl9", "Ccl2", "Tnfsf13b", "Tnfsf11", "Kitl")

levels(seurat)
colVec <- colcond3
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

### avg heatmap cluster marker genes (Cheng et al. Nat. commun. 2022)
```{r avg heatmap SILT FRC cluster marker-2}
Idents(seuratFbwt) <- seuratFbwt$clustername
levels(seuratFbwt)
seurat <- seuratFbwt
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     #ordVec <- c("naive","infectedD4","granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
genesPlot <- c("Clu", "Il7", "Vcam1", "Icam1", "Cxcl16", "Cxcl13", "Cxcl10", "Cxcl9", "Ccl2", "Tnfsf13b", "Tnfsf11", "Kitl")

levels(seurat)
colVec <- colclustername
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

### avg heatmap sel FRC marker genes 
```{r avg heatmap sel FRC marker genes}
Idents(seuratFbwt) <- seuratFbwt$clustername
levels(seuratFbwt)
seurat <- seuratFbwt
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     #ordVec <- c("naive","infectedD4","granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
genesPlot <- c("Ccl19", "Ccl21a", "Ccl2", "Il7","Tnfsf11", "Clu", "Kitl", "Ltbr", "Irf7")

levels(seurat)
colVec <- colclustername
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

### avg heatmap sel FRC marker genes wt only
```{r avg heatmap sel FRC marker genes wt only}
Idents(seuratFbwt) <- seuratFbwt$cond3
levels(seuratFbwt)
seurat <- seuratFbwt
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     #ordVec <- c("naive","infectedD4","granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
genesPlot <- c("Ccl19", "Ccl21a", "Ccl2", "Il7","Tnfsf11", "Clu", "Kitl", "Ltbr", "Irf7")

levels(seurat)
colVec <- colcond3
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

### avg heatmap sel FRC marker genes wt vs cko
```{r avg heatmap sel FRC marker genes wt vs cko}
Idents(seuratFb) <- seuratFb$cond2_plus_cond3
levels(seuratFb)
seurat <- seuratFb
levels(seurat)

avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  #selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(_.*)","",colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=colnames(logNormExpresMa))
  colnames(annotation_col)[1] <- "celltype"
  colnames(annotation_col)[2] <- "cond"
    annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^.*?_)","",cond))

 rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colcond3,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
     ordVec <- c("WT_naive","cko_naive","WT_infectedD4","cko_infectedD4","WT_granulomaD4","cko_granulomaD4")
     
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

##genes
genesPlot <- c("Ccl19", "Ccl21a", "Ccl2", "Il7","Tnfsf11", "Clu", "Kitl", "Ltbr", "Irf7")

levels(seurat)
colVec <- colcond2
# colVec <- c(colPal, colPal, colPal)
# colVec <- c("blue", "red")
avgHeatmap(seurat, selGenes = genesPlot, colVecIdent = colVec)
```

## violin plots of sel FRC genes wt only
```{r violin plots of sel FRC genes wt only}
Idents(seuratFbwt) <- seuratFbwt$cond3
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000071005.Ccl19", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000094686.Ccl21a", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000035385.Ccl2", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000040329.Il7", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000022015.Tnfsf11", pt.size = 1, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000020310.Madcam1", pt.size = 1, cols = colcond3)
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000022037.Clu", pt.size = 1, cols = colcond3)
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000019966.Kitl", pt.size = 0, cols = colcond3)
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000030339.Ltbr", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000025498.Irf7", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000031497.Tnfsf13b", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFbwt, features = "ENSMUSG00000029417.Cxcl9", pt.size = 0, cols = colcond3) 
VlnPlot(object=seuratFb, features = "ENSMUSG00000066551.Hmgb1", pt.size = 0, cols = colcond3, split.by = "cond2") 
```

## dotplots of sel FRC genes wt only
```{r dotplots of sel FRC genes wt only}
genes <- data.frame(gene=rownames(seuratFbwt)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

selGenes <- data.frame(geneID=rev(c("Ccl19", "Ccl21a", "Ccl2", "Il7", "Tnfsf11", "Clu", "Kitl", "Ltbr", "Irf7"))) %>% left_join(., genes, by="geneID")

DotPlot(seuratFbwt, features = selGenes, group.by= "cond3") + RotatedAxis() + scale_color_viridis(option="E") + coord_flip()
DotPlot(seuratFbwt, features = selGenes, group.by= "cond3") + RotatedAxis() + scale_color_viridis(option="E") 
DotPlot(seuratFbwt, features = selGenes, group.by= "cond3") + RotatedAxis() + scale_color_viridis(option="F") + coord_flip()

```

## dotplots of sel FRC genes wt vs cko
```{r dotplots of sel FRC genes wt vs cko}
seuratFb$cond2_plus_cond3 <- factor(seuratFb$cond2_plus_cond3, levels=c("WT_naive","cko_naive","WT_infectedD4","cko_infectedD4","WT_granulomaD4","cko_granulomaD4"))
Idents(seuratFb) <- seuratFb$cond2_plus_cond3
table(seuratFb$cond2_plus_cond3)

genes <- data.frame(gene=rownames(seuratFb)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

selGenes <- data.frame(geneID=rev(c("Ccl19", "Ccl21a", "Ccl2", "Il7", "Tnfsf11", "Madcam1", "Clu", "Kitl", "Ltbr", "Irf7"))) %>% left_join(., genes, by="geneID")

DotPlot(seuratFb, features = selGenes, group.by= "cond2_plus_cond3") + RotatedAxis() + scale_color_viridis(option="E") + coord_flip()
DotPlot(seuratFb, features = selGenes, group.by= "cond2_plus_cond3") + RotatedAxis() + scale_color_viridis(option="F") + coord_flip()
```

## convert to sce 
```{r convert to sce}
##convert seurat object to sce object
sce <- as.SingleCellExperiment(seuratFb)
genes <- data.frame(geneID=rownames(sce)) %>% mutate(gene=gsub(".*\\.", "", geneID))
pal = colorRampPalette(c("#053061", "#2166ac", "#f7f7f7", "#f4a582", "#b2183c", "#85122d"))

##subset naive and convert to sce
seuratFbnaive <- subset(seuratFb, cond3 == "naive")
DimPlot(seuratFbnaive, reduction= "umap") + theme(legend.position = "none")
sceFbnaive <- as.SingleCellExperiment(seuratFbnaive)

##subset SI and convert to sce
seuratFbinf <- subset(seuratFb, cond3 == "infectedD4")
DimPlot(seuratFbinf, reduction= "umap") + theme(legend.position = "none")
sceFbinf <- as.SingleCellExperiment(seuratFbinf)

##subset Gran and convert to sce
seuratFbgran <- subset(seuratFb, cond3 == "granulomaD4")
DimPlot(seuratFbgran, reduction= "umap") + theme(legend.position = "none")
sceFbgran <- as.SingleCellExperiment(seuratFbgran)
```

## umap FRC signature
```{r umap FRC signature}
selFRCgenes <- c("Ccl19", "Ccl21a", "Ccl2", "Il7", "Tnfsf11", "Clu", "Kitl", "Ltbr", "Irf7")

##make a count matrix of signature genes
signGenes <- genes %>% dplyr::filter(gene %in% selFRCgenes)

sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1))
sceSub$sign2[which(sceSub$sign > 1)] <- 1
##check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2") + sc +
  theme(legend.position = "none", point_size = 1)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes naive
sceFbnaiveSub <- sceFbnaive[which(rownames(sceFbnaive) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbnaiveSub@assays@data$logcounts)))/nrow(signGenes)
sceFbnaiveSub$sign <- cntMat
sceFbnaiveSub$sign2 <- sceFbnaiveSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1))
sceFbnaiveSub$sign2[which(sceFbnaiveSub$sign > 1)] <- 1
##check max and min values
max(sceFbnaiveSub$sign)
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes inf
sceFbinfSub <- sceFbinf[which(rownames(sceFbinf) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbinfSub@assays@data$logcounts)))/nrow(signGenes)
sceFbinfSub$sign <- cntMat
sceFbinfSub$sign2 <- sceFbinfSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1))
sceFbinfSub$sign2[which(sceFbinfSub$sign > 1)] <- 1
##check max and min values
max(sceFbinfSub$sign)
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes gran
sceFbgranSub <- sceFbgran[which(rownames(sceFbgran) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbgranSub@assays@data$logcounts)))/nrow(signGenes)
sceFbgranSub$sign <- cntMat
sceFbgranSub$sign2 <- sceFbgranSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1))
sceFbgranSub$sign2[which(sceFbgranSub$sign > 1)] <- 1
##check max and min values
max(sceFbgranSub$sign)
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc
```

## convert to sce wt only
```{r convert to sce wt only}
##convert seurat object to sce object
sce <- as.SingleCellExperiment(seuratFbwt)
genes <- data.frame(geneID=rownames(sce)) %>% mutate(gene=gsub(".*\\.", "", geneID))
pal = colorRampPalette(c("#053061", "#2166ac", "#f7f7f7", "#f4a582", "#b2183c", "#85122d"))

##subset naive and convert to sce
seuratFbnaive <- subset(seuratFbwt, cond3 == "naive")
DimPlot(seuratFbnaive, reduction= "umap") + theme(legend.position = "none")
sceFbnaive <- as.SingleCellExperiment(seuratFbnaive)

##subset SI and convert to sce
seuratFbinf <- subset(seuratFbwt, cond3 == "infectedD4")
DimPlot(seuratFbinf, reduction= "umap") + theme(legend.position = "none")
sceFbinf <- as.SingleCellExperiment(seuratFbinf)

##subset Gran and convert to sce
seuratFbgran <- subset(seuratFbwt, cond3 == "granulomaD4")
DimPlot(seuratFbgran, reduction= "umap") + theme(legend.position = "none")
sceFbgran <- as.SingleCellExperiment(seuratFbgran)
```

## umap FRC signature wt only
```{r umap FRC signature wt only}
selFRCgenes <- c("Ccl19", "Ccl21a", "Ccl2", "Il7", "Tnfsf11", "Clu", "Kitl", "Ltbr", "Irf7")

##make a count matrix of signature genes
signGenes <- genes %>% dplyr::filter(gene %in% selFRCgenes)

sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1))
sceSub$sign2[which(sceSub$sign > 1)] <- 1
##check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2") + sc +
  theme(legend.position = "none", point_size = 1)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes naive
sceFbnaiveSub <- sceFbnaive[which(rownames(sceFbnaive) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbnaiveSub@assays@data$logcounts)))/nrow(signGenes)
sceFbnaiveSub$sign <- cntMat
sceFbnaiveSub$sign2 <- sceFbnaiveSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1))
sceFbnaiveSub$sign2[which(sceFbnaiveSub$sign > 1)] <- 1
##check max and min values
max(sceFbnaiveSub$sign)
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes inf
sceFbinfSub <- sceFbinf[which(rownames(sceFbinf) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbinfSub@assays@data$logcounts)))/nrow(signGenes)
sceFbinfSub$sign <- cntMat
sceFbinfSub$sign2 <- sceFbinfSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1))
sceFbinfSub$sign2[which(sceFbinfSub$sign > 1)] <- 1
##check max and min values
max(sceFbinfSub$sign)
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes gran
sceFbgranSub <- sceFbgran[which(rownames(sceFbgran) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbgranSub@assays@data$logcounts)))/nrow(signGenes)
sceFbgranSub$sign <- cntMat
sceFbgranSub$sign2 <- sceFbgranSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1))
sceFbgranSub$sign2[which(sceFbgranSub$sign > 1)] <- 1
##check max and min values
max(sceFbgranSub$sign)
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc
```

## session info
```{r date and session info}
date()
sessionInfo()
```
